---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

This code converts UN migrant stock data to be used in Gephi for network analysis and visualisation. First, it extracts the UN stock migrants data from excel file. Then, The country names are converted to two-digit codes. Finally, the source-destination data is exported as csv files.

Load packages.

```{r echo=FALSE}
library(readxl)
library(tidyverse)
library(reshape2)
```
## Read data

The data comes fom UN's "International migrant stock: The 2017 revision". The total stock data comes from Table 1 tab in the data "Origin and destination" excel file [http://www.un.org/en/development/desa/population/migration/data/estimates2/data/UN_MigrantStockByOriginAndDestination_2017.xlsx]. It must be noted that the earlier version of the data had different tabs for each year the data was available, whereas the current version has the data for all years in Table 1. 

```{r}
dataloc <- "C:/Users/Lenovo/github/global/data/UN_MigrantStockByOriginAndDestination_2017.xlsx"
clist <- "C:/Users/Lenovo/github/global/data/countriesun.xlsx"
countries <- read_excel(clist, sheet="UN")
data2017 <- read_xlsx(dataloc, sheet = "Table 1", skip = 15, na = "..",
                       col_names = TRUE,
                       col_types = c("numeric", "skip", "text", "skip",
                                     "numeric",
                                     rep("skip", 4),
                                     rep("numeric", 232))) #gives a warning
data2017 <- data2017 %>%
  filter(X__3 < 900)
```
Next, convert the country/place names to two-character codes. The list of all regions are provided in the same excel file. However, the codes have to be linked.

```{r eval=FALSE}
#classifications of countries and areas
countries.un <- read_xlsx(dataloc, sheet = "ANNEX", skip = 15,
                       col_types = c("numeric", "text", "skip", "numeric",
                                     rep("text", 9)))
#filter only countries
countries.un <- countries.un %>%
  filter(Code < 900)
#read codes from comtrade dataset
datacomtrade <- "C:/Users/Lenovo/github/global/data/Comtrade Country Code and ISO list.xlsx"
countries.iso <- read_xlsx(datacomtrade, sheet = 1, skip = 0,
                       col_types = c("numeric", rep("text", 5), "skip", "skip"))
```

```{r eval=FALSE}
data2015 <- read_excel(dataloc, sheet = "Table 16", skip = 15,
                       col_types = c("skip", "text", "skip", "numeric",
                                     rep("skip", 4), rep("numeric", 232)))
```

Functions

```{r}
#* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * *
#*     Functions
#* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * *
readMigrationTable <- function(xyear = 2013) {
  data2015 <- read_excel(dataloc, sheet = "Table 16", skip = 15,
                       col_types = c("skip", "text", "skip", "numeric",
                                     rep("skip", 4), rep("numeric", 232)))
  data2010 <- read_excel(dataloc, sheet = "Table 13", skip = 15,
                       col_types = c("skip", "text", "skip", "numeric",
                                     rep("skip", 4), rep("numeric", 232)))
  data2005 <- read_excel(dataloc, sheet = "Table 10", skip = 15,
                       col_types = c("skip", "text", "skip", "numeric",
                                     rep("skip", 4), rep("numeric", 232)))
  data2000 <- read_excel(dataloc, sheet = "Table 7", skip = 15,
                       col_types = c("skip", "text", "skip", "numeric",
                                     rep("skip", 4), rep("numeric", 232)))
  data1995 <- read_excel(dataloc, sheet = "Table 4", skip = 15,
                       col_types = c("skip", "text", "skip", "numeric",
                                     rep("skip", 4), rep("numeric", 232)))
  data1990 <- read_excel(dataloc, sheet = "Table 1", skip = 15,
                       col_types = c("skip", "text", "skip", "numeric",
                                     rep("skip", 4), rep("numeric", 232)))
  #add datayear column
  data2015$datayear <- 2015
  data2010$datayear <- 2010
  data2005$datayear <- 2005
  data2000$datayear <- 2000
  data1995$datayear <- 1995
  data1990$datayear <- 1990
  data <- rbind(data2015, data2010, data2005, data2000, data1995, data1990)
  return(data)
}
## match with country codes
getCountryCode <- function(xcountry="Nepal") {
  #usage getCountryCode("Australia") will return "AU"
  code <- countries[countries$newname == xcountry,c("ISOCODE")]
  if (is.na(code[1])) {
    return(NA)
  } else {
    return(as.character(code[1]))
  }
}
```
```{r}
getCountryCode <- function(xcountry="Nepal") {
    #usage getCountryCode("Australia") will return "AU"
    code <- countries[countries$newname == xcountry,c("ISOCODE")]
    if (is.na(code[1])) {
      return(NA)
    } else {
      return(as.character(code[1]))
    }
  }
  ## match with country codes
  getCountryName <- function(xcountry="NP") {
    #usage getCountryName("AU") will return "Australia"
    code <- countries[countries$ISOCODE == xcountry, c("COUNTRY_UN")]
    if (is.na(code[1])) {
      return(NA)
    } else {
      return(as.character(code[1]))
    }
  }
```

Clean UN data

```{r}
  # data2015 <- readMigrationTable() #not only for 2013 though
data2015 <- data2017
  # data2015 <- data2015[data2015$X__3 < 900,] #isolate countries only
#process country names
  countries$newname <- chartr("'", " ", countries$COUNTRY_UN)
  countries$newname <- chartr("(", " ", countries$newname)
  countries$newname <- chartr(")", " ", countries$newname)
  countries$newname <- chartr("-", " ", countries$newname)
  countries$newname <- chartr(".", " ", countries$newname)
  countries$newname <- gsub("\\s","", chartr(",", " ", countries$newname))
  #convert col names to country ISCOCODEs
  oldnames <- names(data2015)
  oldnames <- chartr("'", " ", oldnames)
    oldnames <- chartr("(", " ", oldnames)
      oldnames <- chartr(")", " ", oldnames)
        oldnames <- chartr("-", " ", oldnames)
                oldnames <- chartr(",", " ", oldnames)
  newnames <- chartr(".", " ", oldnames) #replace . with space
  newnames <- gsub("\\s","", newnames) #final names to match
  countries$ISOCODE <- as.character(countries$ISOCODE)
  newnames2 <- sapply(newnames[4:length(newnames)], getCountryCode)
  # newnames2[is.na(newnames2)] <- oldnames[is.na(newnames2)]
  names(data2015) <- c("datayear", "name" , "ccode", newnames2)
  data2015$newname <- chartr("'", " ", data2015$name)
  data2015$newname <- chartr("(", " ", data2015$newname)
  data2015$newname <- chartr(")", " ", data2015$newname)
  data2015$newname <- chartr("-", " ", data2015$newname)
  data2015$newname <- chartr(".", " ", data2015$newname)
  data2015$newname <- gsub("\\s","", chartr(",", " ", data2015$newname))
  data2015$ISOCODE <- sapply(data2015$newname, getCountryCode)
  #melt data
  data2015.sub <- data2015[ , c(-2,-3)]
  m2015 <- melt(data2015.sub, c("newname", "ISOCODE", "datayear"),
                names(data2015)[4:235],
                value.name = "STOCK")
  names(m2015) <- c("newname", "source", "year", "destination", "stock")
  m2015 <- m2015[complete.cases(m2015),]
  # m2015 <- m2015[m2015$datayear == 2015,]
  write.csv(m2015, "stock data all.csv", na = "", row.names = FALSE)
  # data2015 <- data2015[data2015$datayear == 2015,] #filter 2015 only
  #sort
  # data2015 <- data2015[order(data2015$ISOCODE), ]
  # write.csv(data2015, "stock matrix 2015.csv", col.names = FALSE, na.)
  # write.table(data2015, "stock matrix 2015b.csv",
  #           na = "",
  #           row.names = FALSE,
  #           # col.names = FALSE,
  #           # append = TRUE,
  #           sep = ",")
  head(countries)
  write.csv(countries[ ,c(1:4)], "nodes.csv", row.names = FALSE)
```

Network graph using igraph

```{r}
nodes.au <- airports.au[,c("V5", "V2")]
names(nodes.au) <- c("id", "airport")
#make links
links.au <- routes.au[, c("V3", "V5", "flights")]
names(links.au) <- c("from", "to", "flights")
library(igraph)
net.au <- graph_from_data_frame(d=links.au, vertices=nodes.au, directed=TRUE) 
# class(net)
plot(net.au, edge.arrow.size=.4)
# net.au.s <- simplify(net.au, remove.multiple = TRUE) 
plot(net.au, edge.arrow.size=.4, edge.curved=.1,
     vertex.label.color="black",
     vertex.color="orange")
#plotting
# V(net.au)$color <- "cyan"
E(net.au)$width <- E(net.au)$flights/2
l <- layout_with_graphopt(net.au)
plot(net.au, edge.arrow.size=.3, edge.curved=0.1,
     edge.color = "darkgreen",
     # vertex.frame.color="#555555",
     # vertex.label=V(net)$media,
     vertex.label.color="black",
     vertex.label.cex=.9, vertex.shape="square",
     layout=l)
#centrality > degree
degree(net.au, mode="in")
centr_degree(net.au, mode="in", normalized=TRUE)
library(RColorBrewer)
pal <- brewer.pal(9, "Paired")
pal2 <- brewer.pal(9, "Blues")
plot(net.au, edge.arrow.size=.3, edge.curved=0.1,
     edge.color = "grey30",
     vertex.color=pal,
     vertex.label.color="black",
     vertex.label.cex=.9, vertex.shape="square",
     layout=l)
#visnetwork
require(visNetwork, quietly = TRUE)
data <- toVisNetworkData(net.au)
visNetwork(nodes = data$nodes, edges = data$edges, height = "700px")
visIgraph(net.au)

visNetwork(nodes.au, links.au, height = "700px") %>%
  visIgraphLayout() %>%
  visNodes(size = 10)
#all layouts
layouts <- grep("^layout_", ls("package:igraph"), value=TRUE)[-1] 

# Remove layouts that do not apply to our graph.

layouts <- layouts[!grepl("bipartite|merge|norm|sugiyama|tree", layouts)]



par(mfrow=c(3,5), mar=c(1,1,1,1))

for (layout in layouts) {
  
  print(layout)
  
  l <- do.call(layout, list(net.au)) 
  
  plot(net.au, edge.arrow.mode=0, layout=l, main=layout) }
#edge colour based on source
tkid <- tkplot(net.au) #tkid is the id of the tkplot that will open

l <- tkplot.getcoords(tkid) # grab the coordinates from tkplot

tk_close(tkid, window.close = T)
par(mfrow=c(1,1))
plot(net.au, layout=l)
#

edge.start <- ends(net.au, es=E(net.au), names=F)[,1]

edge.col <- V(net.au)$color[edge.start]

plot(net, edge.color=edge.col, edge.curved=.1)  

plot(net.au)
```




