---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

This code converts UN migrant stock data to be used in Gephi for network analysis and visualisation. First, it extracts the UN stock migrants data from excel file. Then, The country names are converted to two-digit codes. Finally, the source-destination data is exported as csv files for visualisation in Gephi.

Load packages.

```{r echo=FALSE}
library(readxl)
library(tidyverse)
library(reshape2)
```
## Read data

The data comes fom UN's "International migrant stock: The 2017 revision". The total stock data comes from Table 1 tab in the data "Origin and destination" excel file [http://www.un.org/en/development/desa/population/migration/data/estimates2/data/UN_MigrantStockByOriginAndDestination_2017.xlsx]. It must be noted that the earlier version of the data had different tabs for each year the data was available, whereas the current version has the data for all years in Table 1. 

```{r}
dataloc <- "C:/Users/Lenovo/github/global/data/UN_MigrantStockByOriginAndDestination_2017.xlsx"
clist <- "C:/Users/Lenovo/github/global/data/countriesun.xlsx"
countries <- read_excel(clist, sheet="UN")
data2017 <- read_xlsx(dataloc, sheet = "Table 1", skip = 15, na = "..",
                       col_names = TRUE,
                       col_types = c("numeric", "skip", "text", "skip",
                                     "numeric",
                                     rep("skip", 4),
                                     rep("numeric", 232))) #gives a warning
data2017 <- data2017 %>%
  filter(X__3 < 900)
```
Next, convert the country/place names to two-character codes. The list of all regions are provided in the same excel file. However, the codes have to be linked.

```{r eval=FALSE}
#classifications of countries and areas
countries.un <- read_xlsx(dataloc, sheet = "ANNEX", skip = 15,
                       col_types = c("numeric", "text", "skip", "numeric",
                                     rep("text", 9)))
#filter only countries
countries.un <- countries.un %>%
  filter(Code < 900)
```

Functions

```{r}
## match with country codes
getCountryCode <- function(xcountry="Nepal") {
  #usage getCountryCode("Australia") will return "AU"
  code <- countries[countries$newname == xcountry,c("ISOCODE")]
  if (is.na(code[1])) {
    return(NA)
  } else {
    return(as.character(code[1]))
  }
}
```

```{r}
getCountryCode <- function(xcountry="Nepal") {
    #usage getCountryCode("Australia") will return "AU"
    code <- countries[countries$newname == xcountry,c("ISOCODE")]
    if (is.na(code[1])) {
      return(NA)
    } else {
      return(as.character(code[1]))
    }
  }
  ## match with country codes
  getCountryName <- function(xcountry="NP") {
    #usage getCountryName("AU") will return "Australia"
    code <- countries[countries$ISOCODE == xcountry, c("COUNTRY_UN")]
    if (is.na(code[1])) {
      return(NA)
    } else {
      return(as.character(code[1]))
    }
  }
```

Clean data and export as nodes and edges

```{r}
data2015 <- data2017
#process country names
countries$newname <- chartr("'", " ", countries$COUNTRY_UN)
countries$newname <- chartr("(", " ", countries$newname)
countries$newname <- chartr(")", " ", countries$newname)
countries$newname <- chartr("-", " ", countries$newname)
countries$newname <- chartr(".", " ", countries$newname)
countries$newname <- gsub("\\s","", chartr(",", " ", countries$newname))
#convert col names to country ISCOCODEs
oldnames <- names(data2015)
oldnames <- chartr("'", " ", oldnames)
oldnames <- chartr("(", " ", oldnames)
oldnames <- chartr(")", " ", oldnames)
oldnames <- chartr("-", " ", oldnames)
oldnames <- chartr(",", " ", oldnames)
newnames <- chartr(".", " ", oldnames) #replace . with space
newnames <- gsub("\\s","", newnames) #final names to match
countries$ISOCODE <- as.character(countries$ISOCODE)
newnames2 <- sapply(newnames[4:length(newnames)], getCountryCode)
# newnames2[is.na(newnames2)] <- oldnames[is.na(newnames2)]
names(data2015) <- c("datayear", "name" , "ccode", newnames2)
data2015$newname <- chartr("'", " ", data2015$name)
data2015$newname <- chartr("(", " ", data2015$newname)
data2015$newname <- chartr(")", " ", data2015$newname)
data2015$newname <- chartr("-", " ", data2015$newname)
data2015$newname <- chartr(".", " ", data2015$newname)
data2015$newname <- gsub("\\s","", chartr(",", " ", data2015$newname))
data2015$ISOCODE <- sapply(data2015$newname, getCountryCode)
#melt data
data2015.sub <- data2015[ , c(-2,-3)]
m2015 <- melt(data2015.sub, c("newname", "ISOCODE", "datayear"),
              names(data2015)[4:235],
              value.name = "STOCK")
names(m2015) <- c("newname", "destination", "year", "source", "stock")
m2015 <- m2015[complete.cases(m2015),]
m2015 <- m2015[m2015$year == 2017,]
write.csv(m2015, "edges2017.csv", na = "", row.names = FALSE)
# head(countries)
write.csv(countries[ , c(1:4)], "nodes2017.csv", row.names = FALSE)
```

Network graph using igraph

```{r}
# nodes.au <- airports.au[,c("V5", "V2")]
nodes.mig <- countries[ , c(2:4)]
names(nodes.mig) <- c("id", "lat", "lon")
#make links
# links.au <- routes.au[, c("V3", "V5", "flights")]
links.mig <- m2015[, c("source", "destination", "stock")]
names(links.mig) <- c("source", "target", "weight")
library(igraph)
net.au <- graph_from_data_frame(d = links.mig,
                                vertices = nodes.mig,
                                directed = TRUE)
# class(net)
plot(net.au, edge.arrow.size=.2)
```




