---
title: "R Notebook"
output:
  word_document: default
  html_document:
    df_print: paged
---

# Trade flow matrix

This code reads the BACI: International Trade Database at the Product-Level[http://www.cepii.fr/CEPII/en/publications/wp/abstract.asp?NoDoc=2726] dataset and prepares a matrix of international trade flows for network analysis.

## Load packages

```{r echo=FALSE}
library(tidyverse)
library(data.table)
library(knitr)
```
# Read data

Since the data is very large, we run this only once. For further analysis, we will use the subset for the latest year.

```{r eval=FALSE}
tradedata <- fread('D:/hanh data/trad_charact_0016.csv', header = T, sep = ',')
tradedata <- tradedata %>%
  filter(t == 2016)
saveRDS(tradedata, file = "D:/R/2016data.RDS")
```

Reading prepared data.

```{r echo=FALSE, results='asis'}
tradedata <- readRDS("D:/R/2016data.RDS")
kable(head(tradedata, 12), caption = "First twelve rows of data.")
```
## Data columns

* t: Year of data
* hs6: product HS96 revision
* i: exporter code
* j: importer code
* u: unit value
* trade_type:
* GL:
* price_range:

Summarise by country pairs

```{r echo=FALSE, results='asis'}
#drop year column
tradesum <- tradedata %>% 
  select(-one_of("t")) %>% 
  mutate(pair = paste0(i, "-", j)) %>% 
  group_by(pair) %>% 
  summarise(i = first(i), j = first(j), uvt = sum(uv, na.rm = TRUE)) %>% 
  select(-one_of("pair"))
# nrow(tradesum)
# sum(is.na(tradesum$uvt))
```

```{r}
kable(head(tradesum, 12), caption = "First twelve rows of summarised data.")
```


Save output to disk.

```{r}
write.csv(tradesum, "trade_summary.csv", row.names = FALSE)
```

## Total export by country

```{r}
exportsum <- tradedata %>% 
  select(-one_of("t")) %>% 
  group_by(i) %>% 
  summarise(euvt = sum(uv, na.rm = TRUE)) 
  # head(exportsum)
importsum <- tradedata %>% 
  select(-one_of("t")) %>% 
  group_by(j) %>% 
  summarise(iuvt = sum(uv, na.rm = TRUE))
```


Machine info

```{r}
sessionInfo()
```